\documentclass[12pt]{article}

\usepackage[a4paper,centering,scale=0.8]{geometry}
\usepackage[affil-it]{authblk}

%------------------ImportEnv---------------
\input{../share/macro.tex}
% -----------------ComplexityClass---------
\input{../share/cc.tex}
\def\allfiles{}
\begin{document}
\title{Notes about Algebraic Complexity}
\author{Xinhao Nie}
\affil{Aarhus University}
\maketitle

\section{\SqrtSum}

\begin{Define}
  Given positive integers $a_1, \dots a_n$ and an integer $t,$ we want to decide whether $\sum_{i=1}^n\sqrt{a_i} \leq t.$
\end{Define}


\begin{Lemma}
  We can test whether $\sum_{i=1}^n \sqrt{a_i} = t$ in polynomial
  time.
\end{Lemma}

\begin{Proof}
  The paper 'Decreasing the Nesting Depth of Expressions
  Involving Square Roots' implicitly implies this result in its
  section~5.

  First denest any individual radical that denests, which means for
  all $a_i$ is not perfect sqaure number. Then we choose pair of
  remaining radicals $\sqrt{a_i}$ and $\sqrt{a_j}$ if their product
  denests in $\QQ$. If $\sqrt{a_i} \sqrt{a_j}$ denests as
  $k \in \QQ$, then replace $\sqrt{a_i} + \sqrt{a_j}$ by
  $$
  (1 + \frac{k}{a_i})\sqrt{a_i} \in \QQ(\sqrt{a_i}),
  $$
  and iterate the process of looking for a pair of radicals that
  desnests. Its correctness can be easily checked.

  If at some point no product of a pair of radicals denests, then we
  claim that the entire linear combination could not denest. Next
  we should prove this claim. The rough idea is that we show
  $\forall \sqrt{a_i} = \lambda \ee_j$ where $\lambda \in \QQ$ and
  $\langle \ee_1, \dots,
  \ee_{2^m} \rangle$ span the vector space $\QQ(\sqrt{a_1}, \dots,
  \sqrt{a_m})$ over $\QQ$ if $
  [\QQ(\sqrt{a_1}, \dots, \sqrt{a_n}):\QQ] = 2^m$. Concretely, all
  basis $\ee$ are the subproducts of $m$ square roots. Then, we can
  concluded that if $\sum_{i=1}^n \sqrt{a_i} = t$, each square root
  must be eliminated because they are distinct basis. In other
  words, if there is no product of a pair of radicals denests, then
  it cannot equal to a rational number.

  Together with some fact about basis of extension $\QQ$ over $\QQ$,
  above statement is directly showed by Lemma~1 in the original
  paper. But it's more general, here we beriefly prove
  what we use.
  Suppose that $[\QQ(\sqrt{a_1},\dots,\sqrt{a_m}):\QQ] = 2^m$ and
  prove by induction if
  $\sqrt{a_{m+1}} \in \QQ(\sqrt{a_1},\dots,\sqrt{a_m})$ then
  $\sqrt{a_{m+1}} = \lambda \sqrt{a_1...a_m}$.

  Basis(0): $\lambda = \sqrt{a_1}$.
  Hypothesis(m-1) and Induction(m): Write $\sqrt{a_{m+1}} =
  b\sqrt{a_{m}} + c$, where $b, c \in \QQ(\sqrt{a_1},\dots,
  \sqrt{a_{m-1}})$. By squaring,
  $$
  a_{m+1} = b^2a_m + c^2 + 2bc\sqrt{a_m},
  $$
  which indicates $c = 0$, since $a_{m+1} \in \QQ$. Then, we have
  $$
  \sqrt{a_{m+1}a_m} = ba_{m} \in \QQ(\sqrt{a_1},\dots,\sqrt{a_{m-1}}).
  $$
  By hypothesis, $\sqrt{a_{m+1}a_m} = \lambda \sqrt{a_1...a_{m-1}}$,
  which implies $\sqrt{a_{m+1}} = \lambda a_m^{-1}
  \sqrt{a_1...a_{m-1}a_{m}}$. Thus, we can write each $\sqrt{a_i}
  = \lambda \sqrt{a_1...a_m}$. Moreover, the $2^m$ subproducts of
  $\sqrt{a_1}, \dots, \sqrt{a_m}$ are the basis of $\QQ(\sqrt{a_1},
  \dots,\sqrt{a_n})$ over $\QQ$. \href{https://math.stackexchange.com/questions/30687/the-square-roots-of-different-primes-are-linearly-independent-over-the-field-of}{The proof is here.}

  
\end{Proof}

\end{document}